---
title: "Survey Data Analytics"
author: "Bernard Kilonzo"
date: "2026-06-26"
format: html
code-fold: true
code-summary: "Show the code"
editor: visual
---

## Overview

Survey data analytics helps uncover insights about customer behavior, satisfaction, and demographic patterns. This report demonstrates a **clean**, **reproducible**, and **well‑structured** Quarto workflow for analyzing survey data.

# Load Packages

```{r}
#| label: load-libraries
#| include: true
#| message: false
#| warning: false

library(tidyverse)
library(janitor)
library(DT)
library(gt)
library(scales)
library(paletteer)
```

# Load Data

```{r}
Survey_data <- read.csv("https://raw.githubusercontent.com/bernardkilonzo-rigor/dataviz/main/data/Survey%20Data.csv")%>%
  clean_names()

```

# Shaping Data

```{r}
Survey_data1 <- Survey_data%>%
  pivot_longer(cols = c(q2a:q5_6), names_to = "Quiz", values_to = "Responses")%>%
  pivot_longer(cols = c(q6a:q6e), names_to = "Q6", values_to = "Values")%>%
  mutate(Quiz_group = 
           case_when(
             str_detect(Quiz, "^q2") ~ "Q2",
             str_detect(Quiz, "^q3") ~ "Q3",
             str_detect(Quiz, "^q4") ~ "Q4",
             str_detect(Quiz, "^q5") ~ "Q5",
             TRUE ~ "Others"
           ))
```

# Descriptive Statistics

1.  Gender Frequency

```{r}
Survey_data%>%
  tabyl(gender)%>%
  arrange(desc(n))%>% #sorting before totals
  adorn_totals("row")%>% #add total
  adorn_pct_formatting()%>% #add percentages
  rename(
    Gender = gender,
    Count = n,
    Percent = percent
  )%>%
  gt()
```

2.  Age Group Frequency

```{r}
Survey_data%>%
  tabyl(age_group)%>%
  arrange(desc(n))%>% #sorting before totals
  adorn_totals("row")%>% #add total
  adorn_pct_formatting()%>% #add percentages
  rename(
    "Age Group" = age_group,
    Count = n,
    Percent = percent
  )%>%
  gt()
```

3.  Education Qualifications Frequency

```{r}
Survey_data%>%
  tabyl(highest_qualifications)%>%
  arrange(desc(n))%>% #sorting before totals
  adorn_totals("row")%>% #add total
  adorn_pct_formatting()%>% #add percentages
  rename(
    "Highest Qualifications" = highest_qualifications,
    Count = n,
    Percent = percent
  )%>%
  gt()
```

4.  Employment Status Frequency

```{r}
Survey_data%>%
  tabyl(employment_status)%>%
  arrange(desc(n))%>% #sorting before totals
  adorn_totals("row")%>% #add total
  adorn_pct_formatting()%>% #add percentages
  rename(
    "Employment Status" = employment_status,
    Count = n,
    Percent = percent
  )%>%
  gt()
```

5.  Income Level Frequency

```{r}
Survey_data%>%
  tabyl(income_level)%>%
  arrange(desc(n))%>% #sorting before totals
  adorn_totals("row")%>% #add total
  adorn_pct_formatting()%>% #add percentages
  rename(
    "Income Level" = income_level,
    Count = n,
    Percent = percent
  )%>%
  gt()
```

6.  Country Frequency

```{r}
Survey_data%>%
  tabyl(country)%>%
  arrange(desc(n))%>% #sorting before totals
  adorn_totals("row")%>% #add total
  adorn_pct_formatting()%>% #add percentages
  rename(
    Country = country,
    Count = n,
    Percent = percent
  )%>%
  gt()
```

# Main Analysis

1.  Computing Customer Acquisition by Channel

```{r}
Survey_data1%>%
        group_by(q1)%>%
        summarise(count = n_distinct(respondent_s_id))%>%
        mutate(pr = count/sum(count))%>%
        mutate(Percent = scales::percent(pr, accuracy = 0.1))%>%
        arrange(desc(count))%>%
        mutate(q1 = factor(q1, levels = q1))%>%
        ggplot(aes(y = q1, x = count))+
        geom_bar(stat = "identity", fill = "#FED789")+
        geom_text(aes(label = Percent), hjust = 1)+
        theme(
          panel.background = element_blank(),
          axis.title = element_blank(),
          axis.text = element_text(family = "serif",size = 12, color = "gray30"),
          axis.text.x = element_blank(),
          axis.ticks.x = element_blank()
        )
```

2.  Computing Percentage That Completed the Course.

```{r}
Survey_data1%>%
  filter(Quiz_group == "Q2")%>%
        group_by(Quiz, Responses)%>%
        summarise(count = n_distinct(respondent_s_id))%>%
        mutate(pr = count/sum(count))%>%
        mutate(Percent = formattable::percent(pr, digits = 1))%>%
        ggplot(aes(y = Quiz, x = Percent, fill = Responses))+
        geom_bar(stat = "identity", position = "stack")+
        scale_x_continuous(labels = percent_format())+
        geom_text(aes(label = Percent), position = position_stack(vjust = 0.5),
        color = "white", size =3.5)+
        scale_fill_paletteer_d("nationalparkcolors::Acadia")+
        theme(
          panel.background = element_blank(),
          legend.position = "top",
          legend.title = element_blank(),
          axis.title.y = element_blank()
        )
```

3.  Computing the Most Helpful Content

```{r}
Survey_data1%>%
  filter(Quiz_group == "Q5")%>%
        filter(!is.na(Responses), trimws(Responses) != "")%>%
        group_by(Responses)%>%
        summarise(count = n_distinct(respondent_s_id))%>%
        mutate(pr = count/sum(count))%>%
        mutate(Percent = scales::percent(pr, accuracy = 0.1))%>%
        ggplot(aes(y = Responses, x = count))+
        geom_bar(stat = "identity", fill = "#FED789")+
        geom_text(aes(label = Percent), hjust = 1)+
        theme(
          panel.background = element_blank(),
          axis.title = element_blank(),
          axis.text.x = element_blank(),
          axis.ticks.x = element_blank()
        )
```

4.  Visualizing Level of Satisfaction

```{r}
Survey_data1%>%
  filter(Quiz_group == "Q3")%>%
        mutate(
          Responses = factor(
            Responses,
            levels = c("Highly satisfied", "Satisfied", "Neutral",
                       "Dissatisfied", "Highly dissatisfied")))%>%
        group_by(Quiz, Responses)%>%
        summarise(count = n_distinct(respondent_s_id))%>%
        mutate(pr = count/sum(count))%>%
        mutate(Percent = formattable::percent(pr, digits =1))%>%
        ggplot(aes(y = Quiz, x = Percent, fill = Responses))+
        geom_bar(stat = "identity", position = "stack")+
        scale_x_continuous(labels = percent_format())+
        geom_text(aes(label = Percent), position = position_stack(vjust = 0.5),
                  color = "white", size =3.5)+
        scale_fill_paletteer_d("nationalparkcolors::Acadia")+
        theme(
          panel.background = element_blank(),
          legend.position = "top",
          legend.title = element_blank(),
          legend.text = element_text(family = "sans", size = 7),
          axis.title.y = element_blank()
        )
```

5.  Visualizing Quality of Training Materials

```{r}
Survey_data1%>%
  filter(Quiz_group == "Q4")%>%
        mutate(
          Responses = factor(
            Responses,
            levels = c("Highly satisfied", "Satisfied", "Neutral",
                       "Dissatisfied", "Highly dissatisfied")))%>%
        group_by(Quiz, Responses)%>%
        summarise(count = n_distinct(respondent_s_id))%>%
        mutate(pr = count/sum(count))%>%
        mutate(Percent = formattable::percent(pr, digits = 1))%>%
        ggplot(aes(y = Quiz, x = Percent, fill = Responses))+
        geom_bar(stat = "identity", position = "stack")+
        scale_x_continuous(labels = percent_format())+
        geom_text(aes(label = Percent), position = position_stack(vjust = 0.5),
                  color = "white", size =3.5)+
        scale_fill_paletteer_d("nationalparkcolors::Acadia")+
        theme(
          panel.background = element_blank(),
          legend.position = "top",
          legend.title = element_blank(),
          legend.text = element_text(family = "sans", size = 7),
          axis.title.y = element_blank()
        )
```

6.  Visualizing Net Promoters Score (NPS)

```{r}
Survey_data1%>%
  mutate(nps_group = 
                 case_when(
                   Values >= 9 ~ "Promoters",
                   Values >= 7 ~ "Passive",
                   TRUE ~ "Detractors"
                 ))%>%
        mutate(
          nps_group = factor(
            nps_group,
            levels = c("Promoters", "Passive", "Detractors")
          )
        )%>%
        group_by(Q6, nps_group)%>%
        summarise(count = n_distinct(respondent_s_id))%>%
        mutate(pr = count/sum(count))%>%
        mutate(Percent = formattable::percent(pr, digits = 1))%>%
        ggplot(aes(y = Q6, x = Percent, fill = nps_group))+
        geom_bar(stat = "identity", position = "stack")+
        scale_x_continuous(labels = percent_format())+
        geom_text(aes(label = Percent), position = position_stack(vjust = 0.5),
                  color = "white", size =3.5)+
        scale_fill_paletteer_d("nationalparkcolors::Acadia")+
        theme(
          panel.background = element_blank(),
          legend.position = "top",
          legend.title = element_blank(),
          axis.title.y = element_blank()
        )
```

## Conclusion

Survey data analytics provides a powerful foundation for understanding perceptions, behaviors, and satisfaction levels across different respondent groups. By applying a structured workflow - data cleaning, descriptive summaries, Likert analysis, cross‑tabulations, and visual exploration - we transform raw survey responses into meaningful insights that support evidence‑based decision‑making.

The techniques demonstrated in this document highlight how R enables reproducible, transparent, and scalable analysis. Whether the goal is monitoring customer satisfaction, evaluating service quality, or exploring demographic trends, this workflow ensures that insights remain consistent, defensible, and easy to extend. As organizations continue to rely on data‑driven strategies, robust survey analytics becomes essential for identifying priorities, guiding interventions, and measuring impact over time.
