---
title: "Survey Data Analytics"
author: "Bernard Kilonzo"
date: "2026-06-26"
format: html
editor: visual
---

## Overview

Survey data analytics helps uncover insights about customer behavior, satisfaction, and demographic patterns. This report demonstrates a **clean**, **reproducible**, and **well‑structured** Quarto workflow for analyzing survey data.

# Load Packages

```{r}
#| label: load-libraries
#| include: true
#| message: false
#| warning: false

library(tidyverse)
library(janitor)
library(DT)
```

# Load Data

```{r}
Survey_data <- read.csv("https://raw.githubusercontent.com/bernardkilonzo-rigor/dataviz/main/data/Survey%20Data.csv")|>
  clean_names()

```

## Running Code

When you click the **Render** button a document will be generated that includes both content and the output of embedded code. You can embed code like this:

```{r}
1 + 1
```

You can add options to executable code like this

```{r}
#| echo: false
2 * 2
```

The `echo: false` option disables the printing of code (only output is displayed).
