---
title: "Survey Data Analytics"
author: "Bernard Kilonzo"
date: "2026-06-26"
format: html
code-fold: true
code-summary: "Show the code"
editor: visual
---

## Overview

Survey data analytics helps uncover insights about customer behavior, satisfaction, and demographic patterns. This report demonstrates a **clean**, **reproducible**, and **well‑structured** Quarto workflow for analyzing survey data.

# Load Packages

```{r}
#| label: load-libraries
#| include: true
#| message: false
#| warning: false

library(tidyverse)
library(janitor)
library(DT)
library(gt)
```

# Load Data

```{r}
Survey_data <- read.csv("https://raw.githubusercontent.com/bernardkilonzo-rigor/dataviz/main/data/Survey%20Data.csv")%>%
  clean_names()

```

# Shaping Data

```{r}
survey_data1 <- Survey_data%>%
  pivot_longer(cols = c(q2a:q5_6), names_to = "Quiz", values_to = "Responses")%>%
  pivot_longer(cols = c(q6a:q6e), names_to = "Q6", values_to = "Values")%>%
  mutate(Quiz_group = 
           case_when(
             str_detect(Quiz, "^q2") ~ "Q2",
             str_detect(Quiz, "^q3") ~ "Q3",
             str_detect(Quiz, "^q4") ~ "Q4",
             str_detect(Quiz, "^q5") ~ "Q5",
             TRUE ~ "Others"
           ))
```

# Descriptive Statistics

1.  Gender Frequency

```{r}
Survey_data%>%
  tabyl(gender)%>%
  arrange(desc(n))%>% #sorting before totals
  adorn_totals("row")%>% #add total
  adorn_pct_formatting()%>% #add percentages
  rename(
    Gender = gender,
    Count = n,
    Percent = percent
  )%>%
  gt()
```

2.  Age Group Frequency

```{r}
Survey_data%>%
  tabyl(age_group)%>%
  arrange(desc(n))%>% #sorting before totals
  adorn_totals("row")%>% #add total
  adorn_pct_formatting()%>% #add percentages
  rename(
    "Age Group" = age_group,
    Count = n,
    Percent = percent
  )%>%
  gt()
```

3.  Education Qualifications Frequency

```{r}
Survey_data%>%
  tabyl(highest_qualifications)%>%
  arrange(desc(n))%>% #sorting before totals
  adorn_totals("row")%>% #add total
  adorn_pct_formatting()%>% #add percentages
  rename(
    "Highest Qualifications" = highest_qualifications,
    Count = n,
    Percent = percent
  )%>%
  gt()
```

4.  Employment Status Frequency

```{r}
Survey_data%>%
  tabyl(employment_status)%>%
  arrange(desc(n))%>% #sorting before totals
  adorn_totals("row")%>% #add total
  adorn_pct_formatting()%>% #add percentages
  rename(
    "Employment Status" = employment_status,
    Count = n,
    Percent = percent
  )%>%
  gt()
```

5.  Income Level Frequency

```{r}
Survey_data%>%
  tabyl(income_level)%>%
  arrange(desc(n))%>% #sorting before totals
  adorn_totals("row")%>% #add total
  adorn_pct_formatting()%>% #add percentages
  rename(
    "Income Level" = income_level,
    Count = n,
    Percent = percent
  )%>%
  gt()
```

6.  Country Frequency

```{r}
Survey_data%>%
  tabyl(country)%>%
  arrange(desc(n))%>% #sorting before totals
  adorn_totals("row")%>% #add total
  adorn_pct_formatting()%>% #add percentages
  rename(
    Country = country,
    Count = n,
    Percent = percent
  )%>%
  gt()
```

# Main Analysis

1.  Customer Acquisition by Channel

## Running Code

When you click the **Render** button a document will be generated that includes both content and the output of embedded code. You can embed code like this:

```{r}
1 + 1
```

You can add options to executable code like this

```{r}
#| echo: false
2 * 2
```

The `echo: false` option disables the printing of code (only output is displayed).
