---
title: "Surface Temperature (1950 to 2024)"
author: "Bernard Kilonzo"
format: html
code-fold: true
code-summary: "Show the code"
editor: visual
---

## Surface Temperature in Russia

```{r}
#| include: true
#| message: false
#| warning: false

#loading libraries
library(ggridges)
library(tidyverse)

#loading data
surface_temperature<-read_csv("https://raw.githubusercontent.com/bernardkilonzo-rigor/dataviz/main/data/global%20surface%20temperature.csv")

#pivoting dataset
surface_temperature<-surface_temperature%>%
  pivot_longer(cols = -c(Entity,Code, Month),
               names_to = "Year",
               values_to = "Temp")

#rounding-off values & computing month names
surface_temperature<-surface_temperature%>%
  mutate(tempr = round(Temp,3))%>%
  mutate(mon = month.name[Month])

#ordering months
surface_temperature$mon<-factor(surface_temperature$mon, levels = month.name)

#creating ridgeline plot - for russia
surface_temperature%>%filter(Entity=="Russia")%>%
  ggplot(aes(x = tempr, y = mon, fill = stat(x)))+
  geom_density_ridges_gradient(color = "gray50",linewidth = 0.4)+
  scale_fill_gradient2(low = "#1565c0", mid = "#D3DBE7", high = "#c62828", midpoint = 0)+
  labs(title = "Average Surface Temperature in Russia (1950-2024)",
       subtitle = "Data Source: NASA/GISS",
       caption = "Visualization By: Bernard Kilonzo",
       y ="Month", x = "Average surface temperature (in degrees Celsius)",
       fill = "Temp")+
  theme(panel.grid = element_line(color = "#ececec", linewidth = 0.1),
        panel.background = element_rect(fill = "#fdfbfb"),
        axis.ticks = element_blank(),
        plot.title = element_text(family = "sans",face = "bold"),
        plot.subtitle = element_text(family = "sans",face = "italic"),
        plot.caption = element_text(family = "mono",face = "italic")
  )
```

## Surface Temperature in China

```{r}
#| include: true
#| message: false
#| warning: false

#loading libraries
library(ggridges)
library(tidyverse)

#loading data
surface_temperature<-read_csv("https://raw.githubusercontent.com/bernardkilonzo-rigor/dataviz/main/data/global%20surface%20temperature.csv")

#pivoting dataset
surface_temperature<-surface_temperature%>%
  pivot_longer(cols = -c(Entity,Code, Month),
               names_to = "Year",
               values_to = "Temp")

#rounding-off values & computing month names
surface_temperature<-surface_temperature%>%
  mutate(tempr = round(Temp,3))%>%
  mutate(mon = month.name[Month])

#ordering months
surface_temperature$mon<-factor(surface_temperature$mon, levels = month.name)

#creating ridgeline plot - for russia
surface_temperature%>%filter(Entity=="China")%>%
  ggplot(aes(x = tempr, y = mon, fill = stat(x)))+
  geom_density_ridges_gradient(color = "gray50",linewidth = 0.4)+
  scale_fill_gradient2(low = "#1565c0", mid = "#D3DBE7", high = "#c62828", midpoint = 0)+
  labs(title = "Average Surface Temperature in China (1950-2024)",
       subtitle = "Data Source: NASA/GISS",
       caption = "Visualization By: Bernard Kilonzo",
       y ="Month", x = "Average surface temperature (in degrees Celsius)",
       fill = "Temp")+
  theme(panel.grid = element_line(color = "#ececec", linewidth = 0.1),
        panel.background = element_rect(fill = "#fdfbfb"),
        axis.ticks = element_blank(),
        plot.title = element_text(family = "sans",face = "bold"),
        plot.subtitle = element_text(family = "sans",face = "italic"),
        plot.caption = element_text(family = "mono",face = "italic")
  )
```

## Surface Temperature in Canada

```{r}
#| include: true
#| message: false
#| warning: false

#loading libraries
library(ggridges)
library(tidyverse)

#loading data
surface_temperature<-read_csv("https://raw.githubusercontent.com/bernardkilonzo-rigor/dataviz/main/data/global%20surface%20temperature.csv")

#pivoting dataset
surface_temperature<-surface_temperature%>%
  pivot_longer(cols = -c(Entity,Code, Month),
               names_to = "Year",
               values_to = "Temp")

#rounding-off values & computing month names
surface_temperature<-surface_temperature%>%
  mutate(tempr = round(Temp,3))%>%
  mutate(mon = month.name[Month])

#ordering months
surface_temperature$mon<-factor(surface_temperature$mon, levels = month.name)

#creating ridgeline plot - for russia
surface_temperature%>%filter(Entity=="Canada")%>%
  ggplot(aes(x = tempr, y = mon, fill = stat(x)))+
  geom_density_ridges_gradient(color = "gray50",linewidth = 0.4)+
  scale_fill_gradient2(low = "#1565c0", mid = "#D3DBE7", high = "#c62828", midpoint = 0)+
  labs(title = "Average Surface Temperature in Canada (1950-2024)",
       subtitle = "Data Source: NASA/GISS",
       caption = "Visualization By: Bernard Kilonzo",
       y ="Month", x = "Average surface temperature (in degrees Celsius)",
       fill = "Temp")+
  theme(panel.grid = element_line(color = "#ececec", linewidth = 0.1),
        panel.background = element_rect(fill = "#fdfbfb"),
        axis.ticks = element_blank(),
        plot.title = element_text(family = "sans",face = "bold"),
        plot.subtitle = element_text(family = "sans",face = "italic"),
        plot.caption = element_text(family = "mono",face = "italic")
  )
```

## Surface Temperature in Iran

```{r}
#| include: true
#| message: false
#| warning: false

#loading libraries
library(ggridges)
library(tidyverse)

#loading data
surface_temperature<-read_csv("https://raw.githubusercontent.com/bernardkilonzo-rigor/dataviz/main/data/global%20surface%20temperature.csv")

#pivoting dataset
surface_temperature<-surface_temperature%>%
  pivot_longer(cols = -c(Entity,Code, Month),
               names_to = "Year",
               values_to = "Temp")

#rounding-off values & computing month names
surface_temperature<-surface_temperature%>%
  mutate(tempr = round(Temp,3))%>%
  mutate(mon = month.name[Month])

#ordering months
surface_temperature$mon<-factor(surface_temperature$mon, levels = month.name)

#creating ridgeline plot - for russia
surface_temperature%>%filter(Entity=="Iran")%>%
  ggplot(aes(x = tempr, y = mon, fill = stat(x)))+
  geom_density_ridges_gradient(color = "gray50",linewidth = 0.4)+
  scale_fill_gradient2(low = "#1565c0", mid = "#D3DBE7", high = "#c62828", midpoint = 0)+
  labs(title = "Average Surface Temperature in Iran (1950-2024)",
       subtitle = "Data Source: NASA/GISS",
       caption = "Visualization By: Bernard Kilonzo",
       y ="Month", x = "Average surface temperature (in degrees Celsius)",
       fill = "Temp")+
  theme(panel.grid = element_line(color = "#ececec", linewidth = 0.1),
        panel.background = element_rect(fill = "#fdfbfb"),
        axis.ticks = element_blank(),
        plot.title = element_text(family = "sans",face = "bold"),
        plot.subtitle = element_text(family = "sans",face = "italic"),
        plot.caption = element_text(family = "mono",face = "italic")
  )
```

## Surface Temperature in Mongolia

```{r}
#| include: true
#| message: false
#| warning: false

#loading libraries
library(ggridges)
library(tidyverse)

#loading data
surface_temperature<-read_csv("https://raw.githubusercontent.com/bernardkilonzo-rigor/dataviz/main/data/global%20surface%20temperature.csv")

#pivoting dataset
surface_temperature<-surface_temperature%>%
  pivot_longer(cols = -c(Entity,Code, Month),
               names_to = "Year",
               values_to = "Temp")

#rounding-off values & computing month names
surface_temperature<-surface_temperature%>%
  mutate(tempr = round(Temp,3))%>%
  mutate(mon = month.name[Month])

#ordering months
surface_temperature$mon<-factor(surface_temperature$mon, levels = month.name)

#creating ridgeline plot - for russia
surface_temperature%>%filter(Entity=="Mongolia")%>%
  ggplot(aes(x = tempr, y = mon, fill = stat(x)))+
  geom_density_ridges_gradient(color = "gray50",linewidth = 0.4)+
  scale_fill_gradient2(low = "#1565c0", mid = "#D3DBE7", high = "#c62828", midpoint = 0)+
  labs(title = "Average Surface Temperature in Mongolia (1950-2024)",
       subtitle = "Data Source: NASA/GISS",
       caption = "Visualization By: Bernard Kilonzo",
       y ="Month", x = "Average surface temperature (in degrees Celsius)",
       fill = "Temp")+
  theme(panel.grid = element_line(color = "#ececec", linewidth = 0.1),
        panel.background = element_rect(fill = "#fdfbfb"),
        axis.ticks = element_blank(),
        plot.title = element_text(family = "sans",face = "bold"),
        plot.subtitle = element_text(family = "sans",face = "italic"),
        plot.caption = element_text(family = "mono",face = "italic")
  )
```

## Surface Temperature in India

```{r}
#| include: true
#| message: false
#| warning: false

#loading libraries
library(ggridges)
library(tidyverse)

#loading data
surface_temperature<-read_csv("https://raw.githubusercontent.com/bernardkilonzo-rigor/dataviz/main/data/global%20surface%20temperature.csv")

#pivoting dataset
surface_temperature<-surface_temperature%>%
  pivot_longer(cols = -c(Entity,Code, Month),
               names_to = "Year",
               values_to = "Temp")

#rounding-off values & computing month names
surface_temperature<-surface_temperature%>%
  mutate(tempr = round(Temp,3))%>%
  mutate(mon = month.name[Month])

#ordering months
surface_temperature$mon<-factor(surface_temperature$mon, levels = month.name)

#creating ridgeline plot - for russia
surface_temperature%>%filter(Entity=="India")%>%
  ggplot(aes(x = tempr, y = mon, fill = stat(x)))+
  geom_density_ridges_gradient(color = "gray50",linewidth = 0.4)+
  scale_fill_gradient2(low = "#1565c0", mid = "#D3DBE7", high = "#c62828", midpoint = 0)+
  labs(title = "Average Surface Temperature in India (1950-2024)",
       subtitle = "Data Source: NASA/GISS",
       caption = "Visualization By: Bernard Kilonzo",
       y ="Month", x = "Average surface temperature (in degrees Celsius)",
       fill = "Temp")+
  theme(panel.grid = element_line(color = "#ececec", linewidth = 0.1),
        panel.background = element_rect(fill = "#fdfbfb"),
        axis.ticks = element_blank(),
        plot.title = element_text(family = "sans",face = "bold"),
        plot.subtitle = element_text(family = "sans",face = "italic"),
        plot.caption = element_text(family = "mono",face = "italic")
  )
```

## Surface Temperature in Peru

```{r}
#| include: true
#| message: false
#| warning: false

#loading libraries
library(ggridges)
library(tidyverse)

#loading data
surface_temperature<-read_csv("https://raw.githubusercontent.com/bernardkilonzo-rigor/dataviz/main/data/global%20surface%20temperature.csv")

#pivoting dataset
surface_temperature<-surface_temperature%>%
  pivot_longer(cols = -c(Entity,Code, Month),
               names_to = "Year",
               values_to = "Temp")

#rounding-off values & computing month names
surface_temperature<-surface_temperature%>%
  mutate(tempr = round(Temp,3))%>%
  mutate(mon = month.name[Month])

#ordering months
surface_temperature$mon<-factor(surface_temperature$mon, levels = month.name)

#creating ridgeline plot - for russia
surface_temperature%>%filter(Entity=="Peru")%>%
  ggplot(aes(x = tempr, y = mon, fill = stat(x)))+
  geom_density_ridges_gradient(color = "gray50",linewidth = 0.4)+
  scale_fill_gradient2(low = "#1565c0", mid = "#D3DBE7", high = "#c62828", midpoint = 0)+
  labs(title = "Average Surface Temperature in Peru (1950-2024)",
       subtitle = "Data Source: NASA/GISS",
       caption = "Visualization By: Bernard Kilonzo",
       y ="Month", x = "Average surface temperature (in degrees Celsius)",
       fill = "Temp")+
  theme(panel.grid = element_line(color = "#ececec", linewidth = 0.1),
        panel.background = element_rect(fill = "#fdfbfb"),
        axis.ticks = element_blank(),
        plot.title = element_text(family = "sans",face = "bold"),
        plot.subtitle = element_text(family = "sans",face = "italic"),
        plot.caption = element_text(family = "mono",face = "italic")
  )
```

## Surface Temperature in Cuba

```{r}
#| include: true
#| message: false
#| warning: false

#loading libraries
library(ggridges)
library(tidyverse)

#loading data
surface_temperature<-read_csv("https://raw.githubusercontent.com/bernardkilonzo-rigor/dataviz/main/data/global%20surface%20temperature.csv")

#pivoting dataset
surface_temperature<-surface_temperature%>%
  pivot_longer(cols = -c(Entity,Code, Month),
               names_to = "Year",
               values_to = "Temp")

#rounding-off values & computing month names
surface_temperature<-surface_temperature%>%
  mutate(tempr = round(Temp,3))%>%
  mutate(mon = month.name[Month])

#ordering months
surface_temperature$mon<-factor(surface_temperature$mon, levels = month.name)

#creating ridgeline plot - for russia
surface_temperature%>%filter(Entity=="Cuba")%>%
  ggplot(aes(x = tempr, y = mon, fill = stat(x)))+
  geom_density_ridges_gradient(color = "gray50",linewidth = 0.4)+
  scale_fill_gradient2(low = "#1565c0", mid = "#D3DBE7", high = "#c62828", midpoint = 0)+
  labs(title = "Average Surface Temperature in Cuba (1950-2024)",
       subtitle = "Data Source: NASA/GISS",
       caption = "Visualization By: Bernard Kilonzo",
       y ="Month", x = "Average surface temperature (in degrees Celsius)",
       fill = "Temp")+
  theme(panel.grid = element_line(color = "#ececec", linewidth = 0.1),
        panel.background = element_rect(fill = "#fdfbfb"),
        axis.ticks = element_blank(),
        plot.title = element_text(family = "sans",face = "bold"),
        plot.subtitle = element_text(family = "sans",face = "italic"),
        plot.caption = element_text(family = "mono",face = "italic")
  )
```

## Surface Temperature in Australia

```{r}
#| include: true
#| message: false
#| warning: false

#loading libraries
library(ggridges)
library(tidyverse)

#loading data
surface_temperature<-read_csv("https://raw.githubusercontent.com/bernardkilonzo-rigor/dataviz/main/data/global%20surface%20temperature.csv")

#pivoting dataset
surface_temperature<-surface_temperature%>%
  pivot_longer(cols = -c(Entity,Code, Month),
               names_to = "Year",
               values_to = "Temp")

#rounding-off values & computing month names
surface_temperature<-surface_temperature%>%
  mutate(tempr = round(Temp,3))%>%
  mutate(mon = month.name[Month])

#ordering months
surface_temperature$mon<-factor(surface_temperature$mon, levels = month.name)

#creating ridgeline plot - for russia
surface_temperature%>%filter(Entity=="Australia")%>%
  ggplot(aes(x = tempr, y = mon, fill = stat(x)))+
  geom_density_ridges_gradient(color = "gray50",linewidth = 0.4)+
  scale_fill_gradient2(low = "#1565c0", mid = "#D3DBE7", high = "#c62828", midpoint = 0)+
  labs(title = "Average Surface Temperature in Australia (1950-2024)",
       subtitle = "Data Source: NASA/GISS",
       caption = "Visualization By: Bernard Kilonzo",
       y ="Month", x = "Average surface temperature (in degrees Celsius)",
       fill = "Temp")+
  theme(panel.grid = element_line(color = "#ececec", linewidth = 0.1),
        panel.background = element_rect(fill = "#fdfbfb"),
        axis.ticks = element_blank(),
        plot.title = element_text(family = "sans",face = "bold"),
        plot.subtitle = element_text(family = "sans",face = "italic"),
        plot.caption = element_text(family = "mono",face = "italic")
  )
```

## Surface Temperature in Gabon

```{r}
#| include: true
#| message: false
#| warning: false

#loading libraries
library(ggridges)
library(tidyverse)

#loading data
surface_temperature<-read_csv("https://raw.githubusercontent.com/bernardkilonzo-rigor/dataviz/main/data/global%20surface%20temperature.csv")

#pivoting dataset
surface_temperature<-surface_temperature%>%
  pivot_longer(cols = -c(Entity,Code, Month),
               names_to = "Year",
               values_to = "Temp")

#rounding-off values & computing month names
surface_temperature<-surface_temperature%>%
  mutate(tempr = round(Temp,3))%>%
  mutate(mon = month.name[Month])

#ordering months
surface_temperature$mon<-factor(surface_temperature$mon, levels = month.name)

#creating ridgeline plot - for russia
surface_temperature%>%filter(Entity=="Gabon")%>%
  ggplot(aes(x = tempr, y = mon, fill = stat(x)))+
  geom_density_ridges_gradient(color = "gray50",linewidth = 0.4)+
  scale_fill_gradient2(low = "#1565c0", mid = "#D3DBE7", high = "#c62828", midpoint = 0)+
  labs(title = "Average Surface Temperature in Gabon (1950-2024)",
       subtitle = "Data Source: NASA/GISS",
       caption = "Visualization By: Bernard Kilonzo",
       y ="Month", x = "Average surface temperature (in degrees Celsius)",
       fill = "Temp")+
  theme(panel.grid = element_line(color = "#ececec", linewidth = 0.1),
        panel.background = element_rect(fill = "#fdfbfb"),
        axis.ticks = element_blank(),
        plot.title = element_text(family = "sans",face = "bold"),
        plot.subtitle = element_text(family = "sans",face = "italic"),
        plot.caption = element_text(family = "mono",face = "italic")
  )
```

## Surface Temperature in Iraq

```{r}
#| include: true
#| message: false
#| warning: false

#loading libraries
library(ggridges)
library(tidyverse)

#loading data
surface_temperature<-read_csv("https://raw.githubusercontent.com/bernardkilonzo-rigor/dataviz/main/data/global%20surface%20temperature.csv")

#pivoting dataset
surface_temperature<-surface_temperature%>%
  pivot_longer(cols = -c(Entity,Code, Month),
               names_to = "Year",
               values_to = "Temp")

#rounding-off values & computing month names
surface_temperature<-surface_temperature%>%
  mutate(tempr = round(Temp,3))%>%
  mutate(mon = month.name[Month])

#ordering months
surface_temperature$mon<-factor(surface_temperature$mon, levels = month.name)

#creating ridgeline plot - for russia
surface_temperature%>%filter(Entity=="Iraq")%>%
  ggplot(aes(x = tempr, y = mon, fill = stat(x)))+
  geom_density_ridges_gradient(color = "gray50",linewidth = 0.4)+
  scale_fill_gradient2(low = "#1565c0", mid = "#D3DBE7", high = "#c62828", midpoint = 0)+
  labs(title = "Average Surface Temperature in Iraq (1950-2024)",
       subtitle = "Data Source: NASA/GISS",
       caption = "Visualization By: Bernard Kilonzo",
       y ="Month", x = "Average surface temperature (in degrees Celsius)",
       fill = "Temp")+
  theme(panel.grid = element_line(color = "#ececec", linewidth = 0.1),
        panel.background = element_rect(fill = "#fdfbfb"),
        axis.ticks = element_blank(),
        plot.title = element_text(family = "sans",face = "bold"),
        plot.subtitle = element_text(family = "sans",face = "italic"),
        plot.caption = element_text(family = "mono",face = "italic")
  )
```

## Surface Temperature in Poland

```{r}
#| include: true
#| message: false
#| warning: false

#loading libraries
library(ggridges)
library(tidyverse)

#loading data
surface_temperature<-read_csv("https://raw.githubusercontent.com/bernardkilonzo-rigor/dataviz/main/data/global%20surface%20temperature.csv")

#pivoting dataset
surface_temperature<-surface_temperature%>%
  pivot_longer(cols = -c(Entity,Code, Month),
               names_to = "Year",
               values_to = "Temp")

#rounding-off values & computing month names
surface_temperature<-surface_temperature%>%
  mutate(tempr = round(Temp,3))%>%
  mutate(mon = month.name[Month])

#ordering months
surface_temperature$mon<-factor(surface_temperature$mon, levels = month.name)

#creating ridgeline plot - for russia
surface_temperature%>%filter(Entity=="Poland")%>%
  ggplot(aes(x = tempr, y = mon, fill = stat(x)))+
  geom_density_ridges_gradient(color = "gray50",linewidth = 0.4)+
  scale_fill_gradient2(low = "#1565c0", mid = "#D3DBE7", high = "#c62828", midpoint = 0)+
  labs(title = "Average Surface Temperature in Poland (1950-2024)",
       subtitle = "Data Source: NASA/GISS",
       caption = "Visualization By: Bernard Kilonzo",
       y ="Month", x = "Average surface temperature (in degrees Celsius)",
       fill = "Temp")+
  theme(panel.grid = element_line(color = "#ececec", linewidth = 0.1),
        panel.background = element_rect(fill = "#fdfbfb"),
        axis.ticks = element_blank(),
        plot.title = element_text(family = "sans",face = "bold"),
        plot.subtitle = element_text(family = "sans",face = "italic"),
        plot.caption = element_text(family = "mono",face = "italic")
  )
```

## Surface Temperature in Ukraine

```{r}
#| include: true
#| message: false
#| warning: false

#loading libraries
library(ggridges)
library(tidyverse)

#loading data
surface_temperature<-read_csv("https://raw.githubusercontent.com/bernardkilonzo-rigor/dataviz/main/data/global%20surface%20temperature.csv")

#pivoting dataset
surface_temperature<-surface_temperature%>%
  pivot_longer(cols = -c(Entity,Code, Month),
               names_to = "Year",
               values_to = "Temp")

#rounding-off values & computing month names
surface_temperature<-surface_temperature%>%
  mutate(tempr = round(Temp,3))%>%
  mutate(mon = month.name[Month])

#ordering months
surface_temperature$mon<-factor(surface_temperature$mon, levels = month.name)

#creating ridgeline plot - for russia
surface_temperature%>%filter(Entity=="Ukraine")%>%
  ggplot(aes(x = tempr, y = mon, fill = stat(x)))+
  geom_density_ridges_gradient(color = "gray50",linewidth = 0.4)+
  scale_fill_gradient2(low = "#1565c0", mid = "#D3DBE7", high = "#c62828", midpoint = 0)+
  labs(title = "Average Surface Temperature in Ukraine (1950-2024)",
       subtitle = "Data Source: NASA/GISS",
       caption = "Visualization By: Bernard Kilonzo",
       y ="Month", x = "Average surface temperature (in degrees Celsius)",
       fill = "Temp")+
  theme(panel.grid = element_line(color = "#ececec", linewidth = 0.1),
        panel.background = element_rect(fill = "#fdfbfb"),
        axis.ticks = element_blank(),
        plot.title = element_text(family = "sans",face = "bold"),
        plot.subtitle = element_text(family = "sans",face = "italic"),
        plot.caption = element_text(family = "mono",face = "italic")
  )
```
