---
title: "Report & Dashboard Customization"
author: "Bernard Kilonzo"
date: today
format:
 html:
  theme:
   light: cosmo
   dark: darkly
  theme-toggle: true
  toc: true
  toc-depth: 3
  number-sections: true
  smooth-scroll: true
  code-fold: true
  code-summary: "Show the code"
  title-block-style: manuscript
  title-block-banner: "#023743" #background color
  title-block-banner-color: "#FFFFFF" #text color
  page-layout: full
editor: visual
---

## Introduction

This document demonstrates how you can format your Quarto documents/presentation to make them presentable to your users. Information shared in this document is purely for demonstration purpose and therefore doesn't authoritatively suggest anything about the subject.

## Libraries

This section loads all the required libraries for this analysis.

```{r}
#| message: false
#| warning: false
library(tidyverse)
library(paletteer)
```

## The Data

This section loads Sample - Superstore data set used in this document.

```{r}
#| message: false
#| warning: false

superstore<-read.csv("https://raw.githubusercontent.com/bernardkilonzo-rigor/dataviz/main/data/Sample%20-%20Superstore.csv")

```
