---
title: "Report & Dashboard Customization"
author: "Bernard Kilonzo"
date: today
format:
 html:
  theme:
   light: cosmo
   dark: darkly
  theme-toggle: true
  toc: true
  toc-depth: 3
  number-sections: true
  smooth-scroll: true
  code-fold: true
  code-summary: "Show the code"
  title-block-style: manuscript
  title-block-banner: "#023743" #background color
  title-block-banner-color: "#FFFFFF" #text color
  page-layout: full
editor: visual
---

## Introduction

This document demonstrates how you can format your Quarto documents/presentation to make them presentable to your users. Information shared in this document is purely for demonstration purpose and therefore doesn't authoritatively suggest anything about the subject.

## Libraries

This section loads all the required libraries for this analysis.

```{r}
#| message: false
#| warning: false
library(tidyverse)
library(paletteer)
library(scales)
```

## The Data

This section loads Sample - Superstore data set used in this document.

```{r}
#| message: false
#| warning: false

superstore<-read.csv("https://raw.githubusercontent.com/bernardkilonzo-rigor/dataviz/main/data/Sample%20-%20Superstore.csv")

```

## The Analysis

### A stacked bar chart on Revenue by Region

This section showcases the revenue generated by Month and Region as a stacked bar chart.

```{r}
#| message: false
#| warning: false

#modify the data
#add a column that computes month as a string
superstore<-superstore%>%mutate(month=month(Order.Date, label =TRUE))

#compute sales proportion by region and month
Prtions<-superstore%>%group_by(month, Region)%>%
  summarise(sales = sum(Sales))%>%
  mutate(pr=sales/sum(sales))%>%
  mutate(percent=formattable::percent(pr, digits=1))

#create 100% stacked bar chart
Prtions%>%ggplot(aes(x =month, y = percent, fill =Region))+
  geom_bar(stat = "identity", position = "stack")+
  scale_y_continuous(labels = percent_format())+
    scale_fill_paletteer_d("nationalparkcolors::Acadia")+
  labs(title = "Sales Breakdown by Region",
       caption = "Visualization by: Bernard Kilonzo")+
  geom_text(aes(label = percent),
            position = position_stack(vjust = 0.5),
            color = "#BCBD22",
            size = 3)+
    theme(panel.background = element_blank(),
        axis.line = element_line(color = "gray50"),
        axis.ticks = element_line(color = "gray50"),
        axis.text = element_text(family = "serif", size = 9, color = "gray30"),
        axis.title = element_text(family = "serif", face = "bold", size = 10, color = "gray30"),
        legend.title = element_text(family = "serif", face = "bold", size = 10, color = "gray30"),
        legend.text = element_text(family = "serif", size = 9, color = "gray30"),
        legend.position = "top",
        plot.title = element_text(family = "serif", face = "bold", size = 12, color = "gray30"),
        plot.caption = element_text(family = "serif", face = "italic", size = 8),
  )

```

### A Balloon Plot on Revenue by Region & Sub-Category

```{r}
#| message: false
#| warning: false

#preparing data set
computed_dat<-superstore%>%
  group_by(Region, Sub.Category)%>%
  summarise(sales = sum(Sales))

#creating a balloon plot with ggplot2
computed_dat%>%ggplot(aes(x = Region, y = Sub.Category))+
  geom_point(aes(size = sales, fill = sales), shape = 21, color = "black")+
  scale_size(range = c(2,8))+
  scale_fill_viridis_c() +
  labs(title = "Balloon Plot", caption = "Viz By: Bernard Kilonzo",
       x = "Region",
       y = "Sub-Category", size = "Sales", fill = "Sales")+
  theme_minimal()+
  theme(
    panel.grid = element_line(color = "gray95", linewidth = 0.1),
    plot.caption = element_text(family = "serif", face = "italic", color = "gray50"))

```
